---
title: "p8105_hw2_fir2103"
author: "Farizah Rob"
date: "2022-10-03"
output: github_document
---

```{r load_libraries}
library(tidyverse)
library(readxl)
```


### Problem 1

```{r}
nyc_transit <- read_csv("NYC_Transit_Subway_Entrance_And_Exit_Data.csv", col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>%
  janitor::clean_names() %>% 
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, 
    ada) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

The variables in the dataset are `r names(nyc_transit)`. We have information about the line and station name, along with station latitude and longitudes. The variables route1 through route11 display different lines that go through the station, and some examples  are `r unique(nyc_transit$route1)`. Entrance type has the values `r unique(nyc_transit$entrance_type)`. `vending` refers to if there is any vending machine/store in the station. `ada` is a logical variable implying whether the entry/exit is `ada` compliant. There are `r nrow(nyc_transit)` rows and `r ncol(nyc_transit)` columns. 

To tidy the dataset, I cleaned the variable names (all lowercase), kept the variables (columns) we needed through the `select` function. I changed the `ada` variable to logical from character. 

No, I do not think the dataset is tidy - the `entry` and `vending` variable could also be changed to logical. In addition, the `route` variable could be changed into variables for each line specifically and be a binary variable for each. This might be easier for computation. 

```{r}
#Selecting station name and line to get distinct stations
nyc_transit %>% 
  select(station_name, line) %>% 
  distinct
```

* There are 465 distinct stations. 

```{r}
#ADA compliant stations
nyc_transit %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

* 84 stations are ADA compliant. 

```{r}
#Stations without vending that allow entrance
nyc_transit %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```

* 37.7% of station entrances/exits without vending allow entrance. 

```{r}
#Distinct stations that serve the A train
nyc_transit %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

#ADA compliant stations out of the stations that serve the A train
nyc_transit %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

* 60 stations serve the A train. Out of them, 17 are ADA compliant. 

### Problem 2

**Importing Mr. Trash Wheel dataset and tidying**

```{r, message=FALSE}
mr_trash_wheel <- read_excel("Trash Wheel Collection Data.xlsx", 
                          sheet = 1) %>% 
              janitor::clean_names()

mr_trash_wheel <- mr_trash_wheel %>%
  select(dumpster:homes_powered) %>% 
  mutate(dumpster = as.numeric(dumpster), 
         sports_balls=as.integer(round(sports_balls, digits=0)), 
         year = as.numeric(year)) %>%
  drop_na(dumpster)
```

* To tidy the data, I removed the last 3 columns in the dataset (contained notes). The `dumpster` and `year` variable were changed to numeric, the `sports_balls` variable changed to rounded integer. All `NA` values for dumpster were removed. 


**Importing Professor Trash Wheel data sheet and tidying**

```{r}
prof_trash_wheel <- read_excel("Trash Wheel Collection Data.xlsx", 
                          sheet = 2) %>% janitor::clean_names()

prof_trash_wheel <- prof_trash_wheel %>% 
  mutate(dumpster = as.numeric(dumpster), 
         sports_balls = NA, 
         year = as.numeric(year)) %>%
  drop_na(dumpster) %>% 
  relocate(sports_balls, .after = chip_bags)
```

* To tidy the Professort Trash Wheel dataset, I changed the `dumpster` and `year` variables as numeric; removed all `NA` values of `dumpster`. I added a `sports_balls` variable including `NA` values since this variable does not exist in this dataset - will help us bind rows later. 


**Binding the two data frames together**

```{r}
mr_trash_wheel <- mr_trash_wheel %>% 
  mutate(trash_wheel="Mr Trash Wheel") 

prof_trash_wheel <- prof_trash_wheel %>% 
  mutate(trash_wheel="Professor Trash Wheel") 

#Binded rows together and arranged by descending dumpster number
#Tidied the data further my rounding homes_powered to integers

trash_wheel_df <- bind_rows(mr_trash_wheel, prof_trash_wheel) %>% 
  arrange(dumpster) %>% 
  mutate(homes_powered = round(homes_powered))
```

The resulting dataset has `r nrow(trash_wheel_df)` rows and `r ncol(trash_wheel_df)` columns. `dumpster` variable is the ID variable for the dumpster. `weight` is represented in tonnes and `volume` in cubic tonnes per yard. The variables `r colnames(trash_wheel_df)[7:13]` refers to the litter type. Dates of trash collection are also as variables in the dataset. 

* The total weight of trash collected by Professor Trash Wheel is `r sum(prof_trash_wheel$weight_tons)` tonnes. 

* The total number of sports balls collected by Mr. Trash Wheel in 2020 is `r filter(trash_wheel_df, year==2020 & trash_wheel=="Mr Trash Wheel") %>% pull(sports_balls) %>% sum()`. 

### Problem 3 

**Importing and tidying politicans data**

```{r}
pols_month <- read_csv("./fivethirtyeight_datasets/pols-month.csv") %>%
  janitor::clean_names() %>% 
  separate(mon, c("year", "month", "day"), sep = "-", convert = TRUE) %>% 
  mutate(month=month.abb[month], 
         president=ifelse(prez_gop==1, "gop", 
                            ifelse(prez_dem==1, "dem", NA))) %>% 
  select(-c(prez_dem, prez_gop, day)) %>% 
  arrange(year, month)
```

**Importing and tidying `snp` data**

```{r}
snp <- read_csv("./fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(date, c("month", "day", "year"), sep="/", convert=TRUE) %>%
  mutate(year_pref = ifelse(year<=22, 2000, 1900), 
         year = year_pref+year,
         month=month.abb[as.numeric(month)]) %>% 
  select(year, month, close) %>% 
  arrange(year, month)
```

**Importing and tidying `unemployment` data**

```{r}
unemployment_wide <- read_csv("./fivethirtyeight_datasets/unemployment.csv") 

unemployment_long <- pivot_longer(unemployment_wide, 
    Jan:Dec, 
    names_to = "month",
    values_to = "unemployment"
  ) %>% 
  rename(year=Year)

pols_snp <- left_join(pols_month, snp, by = c("year", "month"))
pols_snp_unemp <- left_join(pols_snp, unemployment_long, by = c("year", "month"))
```

The `pols_month` data has `r nrow(pols_month)` rows and `r ncol(pols_month)` columns - it has information about the numbers of politicians (president, senator, governor, representative) who are democrats or republicans.  The `snp` data has `r nrow(snp)` rows and `r ncol(snp)` columns - it contains information about the Standard & Poor's stock market index, the `close` variable which is the closing values of the index on the given `date`. After transforming the `unemployment` data into long-format (by putting the months into one column and unemployment values as another variable), it has `r nrow(unemployment_long)` rows and `r ncol(unemployment_long)` columns - it gives the percentage of unemployment in each `month` of the associated `year`. 

The merged dataset `pols_snp_unemp` has `r nrow(pols_snp_unemp)` rows and `r ncol(pols_snp_unemp)` columns. The range of years represented are `r range(pols_snp_unemp$year)`. We have unemployment data from year `r unemployment_long[1,"year"]`  and snp data from year `r snp[1,"year"]`. Previous years have `NA` values for these. 
